\vspace{-2cm} Inference on the Effects of Observed Features \\ in Latent Space Models for Networks

### 2018-03-12 manuscript_rearranged.tex
revising simulation

For simplicity and computational efficiency we consider a one-dimensional latent space. That is, we create a single unmeasured variable that can be represented using the Euclidean distance between two nodes in a one-dimensional space. We expect our results to generalize to higher dimensions and other variants of the LSM since the tendency for latent variables to be used in explaining structure that cannot be attributed covariates is universal with respect to both the dimension of the latent space and the form of the LSM. [HERE] To generate an observed covariate that exhibits a controllable degree of colinearity with the latent network structure we follow a four-step process. First, we simulate one-dimensional positions for each node, drawing from a standard normal distribution, and calculate the Euclidean distance (i.e., absolute distance) $\mathbf{d}$ between each pair of positions. Second, we generate a target covariance matrix ($\Sigma$) using the $c$-vine method of \cite{lewandowski2009generating} which generates a uniform draw from the space of positive-definite matrices. The $c$-vine method makes use of a parameter $\eta$, which when equal to $1$ generates draws from uniformly from the space of positive-definite matrices, making the correlation between the latent positions and the measured covariate 0 in expectation. As $\eta$ increases above $1$ draws of $\Sigma$ have increasingly weak correlation\footnote{We show the distribution of correlations under each value of $\eta$ used in our simulation in Figure \ref{fig:vine}.}. Third, given the target covariance matrix ($\Sigma$) among the covariates and distances, $\langle \mathbf{x}, \mathbf{d} \rangle$, we derive the conditional mean vector and covariance, assuming that $\mathbf{x}$ has a normal distribution given $\mathbf{d}$ (see \cite[pp. 116--117]{eaton1983} for the conditional normal derivation).
+Fourth, we simulate $x$ as a normal random variable with the respective conditional means and covariance. Finally we standardize $x$ to have zero mean and unit variance to ensure that the scale of estimation errors is comparable across simulations.

>>>>

For simplicity and computational efficiency we consider a one-dimensional latent space. That is, we create a variable that can be represented using the Euclidean distance between two nodes in a one-dimensional space. We expect our results to generalize to higher dimensions and other variants of the LSM since the tendency for latent variables to be used in explaining structure that cannot be attributed covariates is universal with respect to both the dimension of the latent space and the form of the LSM. To generate an observed covariate that exhibits a controllable degree of correlation with the latent network structure we follow a four-step process. First, we simulate one-dimensional positions for each node, drawing from a standard normal distribution, and calculate the Euclidean distance (i.e., absolute distance) $\mathbf{d}$ between each pair of positions. Second, we generate a target covariance matrix ($\Sigma$) using the $c$-vine method of \cite{lewandowski2009generating} [UNPACK FROM https://www.rdocumentation.org/packages/clusterGeneration/versions/1.3.4/topics/genPositiveDefMat] which generates a uniform draw from the space of positive-definite matrices. The $c$-vine method makes use of a parameter $\eta$, which when equal to $1$ generates draws from uniformly from the space of positive-definite matrices, making the correlation between the latent positions and the measured covariate 0 in expectation. As $\eta$ increases above $1$ draws of $\Sigma$ have increasingly weak correlation\footnote{We show the distribution of correlations under each value of $\eta$ used in our simulation in Figure \ref{fig:vine}.}. Third, given the target covariance matrix ($\Sigma$) among the covariates and distances, $\langle \mathbf{x}, \mathbf{d} \rangle$, we derive the conditional mean vector and covariance, assuming that $\mathbf{x}$ has a normal distribution given $\mathbf{d}$ (see \cite[pp. 116--117]{eaton1983} for the conditional normal derivation).
+Fourth, we simulate $x$ as a normal random variable with the respective conditional means and covariance. Finally we standardize $x$ to have zero mean and unit variance to ensure that the scale of estimation errors is comparable across simulations.
